---
title: "DA5020 - Homework 5: Dates and Times"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  # mute messages output
  message = FALSE
)

library(stringr)
library(lubridate)
library(tidyverse)
```

Continue working with Farmers Market data from last week.

This week's assignment is not only about dates and times, but also what you learnt from past weeks: data transformation, strings, and more.

You may also need to go through a review on [R control statesments](http://uc-r.github.io/control_statements) since they will come handy in solving some of the problems.
```{r}
Farmers_Market <- read.csv("farmers_market.csv",header=TRUE,stringsAsFactors = FALSE)

```
## Questions

1. (10 points) Add a new column `Season1Days` that contains the number of days a market is opened per week (for the dates it is open).
```{r}
Days_open <- str_extract_all(Farmers_Market$Season1Time,"\\w\\w\\w:") # extract the days from season1 time column
Farmers_Market$Season1Days <- str_count(Days_open,"\\w\\w\\w:") # count the number of extracting in each row and that gives the number of days a market is open in season 1
Farmers_Market$Season1Days
```
2. (10 points) Add a new column `WeekendOpen` indicating whether a market opens during weekends in `Season1`.
```{r}
#create empty column "WeekendOpen"
Farmers_Market$WeekendOpen <- 0
Seas1_Days <- grep("Sat|Sun",Farmers_Market$Season1Time,ignore.case = TRUE) # extract the row numbers which have saturday or sunday in it using grep and use those row numbers to indicate whether a market is open on weekends or not.
Farmers_Market$WeekendOpen[Seas1_Days] <- "Open on Weekends"
Farmers_Market$WeekendOpen[-Seas1_Days] <- "Closed on Weekends"
Farmers_Market$WeekendOpen
```

3. (20 points) Find out which markets close before 6PM, and which open only for fewer than 4 hours a day. For simplicity, consider only `Season1Time`. For markets with different open hours across a week, use the average length of open hours for the days they actually open. 
```{r}
Seas1_Time <- str_extract_all(Farmers_Market$Season1Time,"-\\s?\\d+:\\d\\d \\w\\w") #extracting closing time by specifying the digits coming after "-"
Seas1_Time_ext <- str_extract_all(Seas1_Time,"\\d+:\\d\\d \\w\\w",simplify = TRUE) #removing the '-'
# converting to 24 hour format
for (k in 1:ncol(Seas1_Time_ext)) {
Seas1_Time_ext[,k] <- format(strptime(Seas1_Time_ext[,k], "%I:%M %p"), format="%H:%M:%S")
}
# replacing "NA" with empty space
Seas1_Time_ext <- ifelse (is.na(Seas1_Time_ext),"",Seas1_Time_ext)
#removing the minutes and seconds part because in order to find if the store closes before 6pm we need to see if the hour part is less than 18 or not
for (k in 1:ncol(Seas1_Time_ext)) {
Seas1_Time_ext[,k] <- str_replace_all(Seas1_Time_ext[,k],":\\d\\d:\\d\\d","")
}
# creating empty matrix to collect the data. this has more than one column because some of the markets have more than one business hours
Seas1_Time_close <- matrix( data = NA ,nrow = nrow(Seas1_Time_ext), ncol = ncol(Seas1_Time_ext))
for (k in 1:ncol(Seas1_Time_ext)) {
  for ( i in 1:nrow(Seas1_Time_ext)){
    if(Seas1_Time_ext[i,k] != "" && Seas1_Time_ext[i,k]<18) {
    Seas1_Time_close[i,k] <- "closes before 6 pm"
    }
  }
}

Seas1_Time_close <- paste(Seas1_Time_close[,1],Seas1_Time_close[,2],Seas1_Time_close[,3],Seas1_Time_close[,4],Seas1_Time_close[,5],Seas1_Time_close[,6],Seas1_Time_close[,7],sep = ",")

Farmers_Market$MarketName[str_which(Seas1_Time_close,"closes before 6 pm")]

# Open less than 4 hrs
Seas1_Time_2 <- str_extract_all(Farmers_Market$Season1Time,"\\d+:\\d\\d \\w\\w\\s?-\\s?\\d+:\\d\\d \\w\\w",simplify = TRUE) # extracting all the business hours
# creating empty matrix
Seas1_openhrs <- matrix( data = NA ,nrow = nrow(Seas1_Time_2), ncol = ncol(Seas1_Time_2))
# go into each column to work on each business hour duration
for (c in 1:ncol(Seas1_Time_2)) {
  Seas1_Time_open <- matrix( data = NA ,nrow = nrow(Seas1_Time_2), ncol = 2)
  Seas1_Time_open <- str_extract_all(Seas1_Time_2[,c],"\\d+:\\d\\d \\w\\w",simplify = TRUE)
 # Seas1_Time_open <- format(strptime(Seas1_Time_open, "%I:%M %p"), format="%H:%M:%S")
  Time1 <- as.POSIXct(format(strptime(Seas1_Time_open[,1],"%I:%M %p"),format="%H:%M:%S"),format = "%H:%M")
  Time2 <- as.POSIXct(format(strptime(Seas1_Time_open[,2],"%I:%M %p"),format="%H:%M:%S"),format = "%H:%M")
Seas1_openhrs[,c] <-str_extract(difftime(Time2,Time1,units = "hours"),"\\d+") # calculation the duration of each businees hours for each market
}
Seas1_meanopen <- NA
for ( i in 1:nrow(Seas1_openhrs)) {
  Seas1_meanopen[i] <- mean(as.numeric(Seas1_openhrs[i,]), na.rm = TRUE) # mean of all the business hours for each market
}
Farmers_Market$MarketName[which(Seas1_meanopen < 4)]


```
3. (40 Points) The seasons are not standardized and would make analysis difficult. Create four new columns for four seasons (Spring, Summer, Fall, Winter), indicating whether a market is available in that season. Also, create two additional columns `HalfYear` and `YearRound` to identify those who open across seasons. Define "half year" and "year round" on your own terms, but explain them before you write the code (or as comments in your code). (Hint: you may want to create even more auxiliary columns, `Season1BeginDate` and `Season1EndDate` for example.)
```{r}
# get begin and end dates for each season
Season1 <- separate(Farmers_Market,Season1Date, into = c("BeginDate","EndDate"), sep = " to ")
Season2 <- separate(Farmers_Market,Season2Date, into = c("BeginDate","EndDate"), sep = " to ")
Season3 <- separate(Farmers_Market,Season3Date, into = c("BeginDate","EndDate"), sep = " to ")
Season4 <- separate(Farmers_Market,Season4Date, into = c("BeginDate","EndDate"), sep = " to ")

# function to convert any abbreviated months to numerical months
dateFun <- function(x, Day = "29", Year = "2016"){
  #using year 2016 since it is a leap year so 29th can be used for february and any dates with 29th february in the data will be handled without any error
  x <- ifelse(substr(x, 1, 3) %in% month.abb,
              paste(match(substr(x, 1, 3), month.abb),
                    Day,
                    Year, sep = "/"), x)
  #return date
  mdy(x)
}
# apply the function to all dates extracted above
Season1$BeginDate <- lapply(Season1$BeginDate, dateFun)
Season1$EndDate <- lapply(Season1$EndDate, dateFun)
Season2$BeginDate <- lapply(Season2$BeginDate, dateFun)
Season2$EndDate <- lapply(Season2$EndDate, dateFun)
Season3$BeginDate <- lapply(Season3$BeginDate, dateFun)
Season3$EndDate <- lapply(Season3$EndDate, dateFun)
Season4$BeginDate <- lapply(Season4$BeginDate, dateFun)
Season4$EndDate <- lapply(Season4$EndDate, dateFun)

# function to define the seasons of begin and end dates
getSeasons <- function(DATES) {
    WS <- as.Date("2015-12-21", format = "%Y-%m-%d") # Winter Solstice
    SE <- as.Date("2016-3-20",  format = "%Y-%m-%d") # Spring Equinox
    SS <- as.Date("2016-6-20",  format = "%Y-%m-%d") # Summer Solstice
    FE <- as.Date("2016-9-22",  format = "%Y-%m-%d") # Fall Equinox

    # Convert dates from any year to 2016 dates
    d<- DATES
    for (i in 1:nrow(Season1)) {
d[[i]]<-as.Date(strftime(DATES[[i]],format="2016-%m-%d"))
}

    ifelse (d >= WS & d < SE, "Winter",
      ifelse (d >= SE & d < SS, "Spring",
        ifelse (d >= SS & d < FE, "Summer","Fall")))
}
# create empty begin and end columns for each season and then apply the function to get the seasons for each market for the begin and end date
Season1$Begin <- 0
Season1$End <- 0
Season2$Begin <- 0
Season2$End <- 0
Season3$Begin <- 0
Season3$End <- 0
Season4$Begin <- 0
Season4$End <- 0

Season1$Begin <- getSeasons(Season1$BeginDate)
Season1$End <- getSeasons(Season1$EndDate)
Season2$Begin <- getSeasons(Season2$BeginDate)
Season2$End <- getSeasons(Season2$EndDate)
Season3$Begin <- getSeasons(Season3$BeginDate)
Season3$End <- getSeasons(Season3$EndDate)
Season4$Begin <- getSeasons(Season4$BeginDate)
Season4$End <- getSeasons(Season4$EndDate)
# create new column for each season saying unavailable which will be updated later according to the dates the market are open for
Farmers_Market$Winter <- "Unavailable"
Farmers_Market$Spring <- "Unavailable"
Farmers_Market$Summer <- "Unavailable"
Farmers_Market$Fall <- "Unavailable"

# go through each season begin and end columns and compare the beginning and ending seasons and accordingly assign the availability for each season.
for (h in 1:nrow(Farmers_Market)) {
  if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Winter" && !is.na(Season1$End[h]) && Season1$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    }
    else if ( !is.na(Season1$Begin[h]) && Season1$Begin[h] == "Winter" && !is.na(Season1$End[h]) && Season1$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Winter" && !is.na(Season1$End[h]) && Season1$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Winter" && !is.na(Season1$End[h]) && Season1$End[h] == "Fall") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Spring" && !is.na(Season1$End[h]) && Season1$End[h] == "Spring") {
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Spring" && !is.na(Season1$End[h]) && Season1$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Spring" && !is.na(Season1$End[h]) && Season1$End[h] == "Summer") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Spring" && !is.na(Season1$End[h]) && Season1$End[h] == "Fall") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Summer") {
    if(!is.na(Season1$End[h]) && Season1$End[h] == "Summer") {
    Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
       Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Fall") {
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season1$Begin[h]) && Season1$Begin[h] == "Fall") {
    if(!is.na(Season1$End[h]) && Season1$End[h] == "Fall") {
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season1$End[h]) && Season1$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
}
for (h in 1:nrow(Farmers_Market)) {
  if (!is.na(Season2$Begin[h]) && Season2$Begin[h] == "Winter") {
    if(!is.na(Season2$End[h]) && Season2$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Fall") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season2$Begin[h]) && Season2$Begin[h] == "Spring") {
    if(!is.na(Season2$End[h]) && Season2$End[h] == "Spring") {
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
    Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Summer") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Fall") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season2$Begin[h]) && Season2$Begin[h] == "Summer") {
    if(!is.na(Season2$End[h]) && Season2$End[h] == "Summer") {
    Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
      Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Fall") {
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season2$Begin[h]) && Season2$Begin[h] == "Fall") {
    if(!is.na(Season2$End[h]) && Season2$End[h] == "Fall") {
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season2$End[h]) && Season2$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
}
for (h in 1:nrow(Farmers_Market)) {
  if (!is.na(Season3$Begin[h]) && Season3$Begin[h] == "Winter") {
    if(!is.na(Season3$End[h]) && Season3$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Fall") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season3$Begin[h]) && Season3$Begin[h] == "Spring") {
    if(!is.na(Season3$End[h]) && Season3$End[h] == "Spring") {
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
       Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Summer") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Fall") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season3$Begin[h]) && Season3$Begin[h] == "Summer") {
    if(!is.na(Season3$End[h]) && Season3$End[h] == "Summer") {
    Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
        Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Fall") {
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
  else if (!is.na(Season3$Begin[h]) && Season3$Begin[h] == "Fall") {
    if(!is.na(Season3$End[h]) && Season3$End[h] == "Fall") {
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season3$End[h]) && Season3$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }
}
for (h in 1:nrow(Farmers_Market)) {
  if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Winter" && !is.na(Season4$End[h]) && Season4$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Winter" && !is.na(Season4$End[h]) && Season4$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Winter" && !is.na(Season4$End[h]) && Season4$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Winter" && !is.na(Season4$End[h]) && Season4$End[h] == "Fall") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Spring" && !is.na(Season4$End[h]) && Season4$End[h] == "Spring") {
    Farmers_Market$Spring[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Spring" && !is.na(Season4$End[h]) && Season4$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
      Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Spring" && !is.na(Season4$End[h]) && Season4$End[h] == "Summer") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Spring" && !is.na(Season4$End[h]) && Season4$End[h] == "Fall") {
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Summer" && !is.na(Season4$End[h]) && Season4$End[h] == "Summer") {
    Farmers_Market$Summer[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Summer" && !is.na(Season4$End[h]) && Season4$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Summer[h] <- "Available"
       Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Summer" && !is.na(Season4$End[h]) && Season4$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
        Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Summer" && !is.na(Season4$End[h]) && Season4$End[h] == "Fall") {
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Fall" && !is.na(Season4$End[h]) && Season4$End[h] == "Fall") {
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Fall" && !is.na(Season4$End[h]) && Season4$End[h] == "Winter") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Fall" && !is.na(Season4$End[h]) && Season4$End[h] == "Spring") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
    else if (!is.na(Season4$Begin[h]) && Season4$Begin[h] == "Fall" && !is.na(Season4$End[h]) && Season4$End[h] == "Summer") {
    Farmers_Market$Winter[h] <- "Available"
    Farmers_Market$Spring[h] <- "Available"
     Farmers_Market$Summer[h] <- "Available"
     Farmers_Market$Fall[h] <- "Available"
    }
  }

Farmers_Market$Winter 
Farmers_Market$Spring 
Farmers_Market$Summer
Farmers_Market$Fall
# creat halfyear and yearround columns indicating "No" which will be updated as per the criteria shown below
Farmers_Market$HalfYear <- "No"  # update to yes If open for 1 to 3 seasons 
Farmers_Market$YearRound <- "No" # update to yes If open for all 4 seasons

for(j in 1:nrow(Farmers_Market)){
  num_Seasons <- sum(str_count(Farmers_Market[j,],"Available"))
  if (!is.na(num_Seasons) && num_Seasons == 4) {
    Farmers_Market$YearRound[j] <- "Yes"
  }
  else if (!is.na(num_Seasons) && num_Seasons >= 1 && num_Seasons < 4){
    Farmers_Market$HalfYear[j] <- "Yes"
  }
}
Farmers_Market$YearRound
Farmers_Market$HalfYear
```
4. (20 points) _Open question_: explore the new variables you just created. Aggregate them at different geographic levels, or some other categorical variable. What can you discover?
```{r}
ggplot(Farmers_Market, aes(x = x, y = y,fill = Winter)) + geom_polygon(color = "white") + theme_bw()
ggplot(Farmers_Market, aes(x = x, y = y,fill = Spring)) + geom_polygon(color = "white") + theme_bw()
ggplot(Farmers_Market, aes(x = x, y = y,fill = Summer)) + geom_polygon(color = "white") + theme_bw()
ggplot(Farmers_Market, aes(x = x, y = y,fill = Fall)) + geom_polygon(color = "white") + theme_bw()

```
By Plotting all 4 seasons according to the longitude and latitude of the markets, we can see that during winter only the markets in the south, southwest and southeast regions are available, in spring markets in all the regions are available, in summer most of the markets available are located in the north,northeast and southeast regions whereas in fall markets in south, southwest, southeast and some markets in northeast are available.

## Submission
You need to submit an .Rmd extension file as well as the generated pdf file. Be sure to state all the assumptions and give explanations as comments in the .Rmd file wherever needed to help us assess your submission. Please name the submission file LAST_FirstInitial_1.Rmd for example for John Smith's 1st assignment, the file should be named Smith_J_1.Rmd. 

